---
title: "Untitled"
output: html_document
date: "2023-03-17"
---

```{r}
library(reshape2)
library(ggplot2)
plot_dev_explained <- function(file_path,aic_path) {
  # Load data
  DevExpl <- readRDS(file_path)
  
  # DETERMINE SIGNIFICANCE WITH AIC
  AICs <- readRDS(aic_path)
  
  # TEMP
  DevExpl=DevExpl[1:5000,]
  AICs=AICs[1:5000,] 
  
  # Calculate deviance explained by each term
  devExplEachTerm <- matrix(
    data = NA,
    nrow = 5000,
    ncol = length(colnames(DevExpl))
  )
  for (r in 2:length(colnames(DevExpl))) {
    columnOfInt <- DevExpl[, r]
    propDev <- 1 - (columnOfInt / DevExpl$devExplained_full)
    devExplEachTerm[, r] <- propDev
  }
  
  # is it "significant"? (more than 99% of runs)
  SigEachTerm <- matrix(
    data = NA,
    nrow = 5000,
    ncol = length(colnames(AICs))
  )
  for (r in 2:length(colnames(AICs))) {
    columnOfInt <- AICs[, r]
    propSig <- (5000-sum((AICs[, 1]<(AICs[, r]))))/5000
    SigEachTerm[, r] <- propSig
  }
  
  # Remove prefix for plotting
  columnname <- colnames(DevExpl)
  columnname <- gsub('devExplained_', '', columnname)
  colnames(devExplEachTerm) <- columnname
  colnames(SigEachTerm)<-columnname
  
  # Melt data for plotting
  plotdf <- melt(data.frame(devExplEachTerm))
  # aic df
  aicdf <- melt(data.frame(SigEachTerm))
  colnames(aicdf)<-c('variable','Sig')
  plotdf<-cbind(plotdf,aicdf)

  # and variable category labels
  categories=c('No Association','Events','Performance','Demographic','Body')
  plotdf$cat<-rep('No Association',dim(plotdf)[1])
  plotdf$cat<-factor(plotdf$cat,levels=categories)
  # Events
  plotdf$cat[plotdf$variable==c('n_died')]<-'Events'
  plotdf$cat[plotdf$variable==c('n_injured')]='Events'
  plotdf$cat[plotdf$variable==c('n_crime')]='Events'
  plotdf$cat[plotdf$variable==c('n_friend')]='Events'
  plotdf$cat[plotdf$variable==c('n_friend_injur')]='Events'
  plotdf$cat[plotdf$variable==c('n_arrest')]='Events'
  plotdf$cat[plotdf$variable==c('n_friend_died')]='Events'
  plotdf$cat[plotdf$variable==c('n_mh')]='Events'
  plotdf$cat[plotdf$variable==c('n_sib')]='Events'
  plotdf$cat[plotdf$variable==c('n_victim')]='Events'
  plotdf$cat[plotdf$variable==c('n_separ')]='Events'
  plotdf$cat[plotdf$variable==c('n_law')]='Events'
  plotdf$cat[plotdf$variable==c('n_school')]='Events'
  plotdf$cat[plotdf$variable==c('n_move')]='Events'
  plotdf$cat[plotdf$variable==c('n_jail')]='Events'
  plotdf$cat[plotdf$variable==c('n_step')]='Events'
  plotdf$cat[plotdf$variable==c('n_new_job')]='Events'
  plotdf$cat[plotdf$variable==c('n_new_sib')]='Events'

  # Performance
  plotdf$cat[plotdf$variable==c('n_g')]='Performance'
  plotdf$cat[plotdf$variable==c('n_Grades')]='Performance'
  
  # Demographics
  plotdf$cat[plotdf$variable==c('n_interview_age')]='Demographic'
  plotdf$cat[plotdf$variable==c('n_income')]='Demographic'
  plotdf$cat[plotdf$variable==c('n_parental_education')]='Demographic'
  plotdf$cat[plotdf$variable==c('n_sex')]='Demographic'
  plotdf$cat[plotdf$variable==c('n_race_ethnicity')]='Demographic'
  
  # Body
  plotdf$cat[plotdf$variable==c('n_weight')]='Body'
  plotdf$cat[plotdf$variable==c('n_waist')]='Body'
  plotdf$cat[plotdf$variable==c('n_height')]='Body'
  plotdf$cat[plotdf$variable==c('n_BMI')]='Body'
  
  # Define custom color scheme for the categories
  colors <- c("Events" = "#FFD700", "Performance" = "#EA4335", "Demographic" = "#4285F4", "Body" = "#34A853", "No Association" = "grey")
  
  # Get levels of factor
  fct_levels <- levels(plotdf$variable)
  # Reorder levels with n_interview_age after n_Grades
  new_levels <- c(fct_levels[1:20], fct_levels[22], fct_levels[21], fct_levels[23:length(fct_levels)])
  # and n_ removal
  new_levels<- gsub("n_", "", new_levels)
  new_variable_names=gsub("n_", "", plotdf$variable)
  plotdf$variable <-factor(new_variable_names,levels=new_levels)
  # last take-outs
  plotdf=subset(plotdf, variable != "full")
  plotdf=subset(plotdf, variable != "parentPcount")
  
  
  # take out events for now: will plot std betas instead
  plotdf=plotdf[plotdf$cat!='Events',]
  
  # set cat insignificant to null color
  plotdf$cat[plotdf$Sig>0.01]='No Association'
  
 # Plot deviance explained by each term with grouping and coloration by 'cat'
  ggplot(data = plotdf, aes(x = factor(variable,levels=rev(c('g','Grades','interview_age','sex','income','race_ethnicity','parental_education','height','weight','waist','BMI'))), y = value,color=cat))  +
  geom_boxplot(width = 0.2, position = position_nudge(x = 0.5)) +
  geom_jitter(aes(alpha = 0.01), size = 3, width = 0.2, height = 0.01) +
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  ylab("Value") +
  xlab("Covariate") +
  theme_minimal(base_size = 70) +
  ylim(c(-0.01, 0.15)) +
  labs(color = "Category")
  
  
  
  
  # ggplot(data = plotdf, aes(x = factor(variable, levels = new_levels), y = value, fill = cat))  +
  # geom_boxplot(aes(fill=cat)) +
  # scale_color_manual(values = colors) +
  # #scale_fill_manual(values = colors) +
  # ylab(label) +
  # xlab('Covariate') +
  # theme_minimal(base_size = 35) +
  # theme(axis.text.x = element_text(angle = 45)) +
  #   ylim(c(-0.005,.14))+
  #   labs(fill = "Category") 
}

```

```{r}
plot_betas <- function(file_path,Signifvec) {
  # Load data
  betas <- readRDS(file_path)
 
  # TEMP
  betas=betas[1:5000,]
  
  ## DETERMINE SIGNIFICANCE WITH AIC
  #AICs <- readRDS(aic_path)
  #AICs=AICs[1:5000,] 
  ## match AIC columns to betas in this instance
  ## YLES are 2:19, income is 24, sex is 26, race is 27
  #AICs=AICs[,c(2:19,24,26,27)]
  ## and apply race AIC to all race factor levels
  #AICs$AIC_n_race_ethnicity_2=AICs[,21]
  #AICs$AIC_n_race_ethnicity_3=AICs[,21]
  #AICs$AIC_n_race_ethnicity_4=AICs[,21]

  
  ## is it "significant"? (more than 99% of runs)
  #SigEachTerm <- matrix(
  #  data = NA,
  #  nrow = 5000,
  #  ncol = length(colnames(AICs))
  #)
  #for (r in 2:length(colnames(AICs))) {
  #  columnOfInt <- AICs[, r]
  #  propSig <- (5000-sum((AICs[, 1]<(AICs[, r]))))/5000
  #  SigEachTerm[, r] <- propSig
  #}
  
  # if it's a temporal precedence file, add catch 
  if (length(grep('TempPrec',file_path))>0){
    # nest to detect inter-vist events
    if (grep('IVE',filepath)==1){
    # Remove prefix for plotting
    columnname <- colnames(betas)
    columnname <- gsub('ple_', '', columnname)
    columnname <- gsub('_y_IVE.y', '', columnname)
    colnames(betas) <- columnname
    }else{
    # Remove prefix for plotting
    columnname <- colnames(betas)
    columnname <- gsub('ple_', '', columnname)
    columnname <- gsub('_y.x', '', columnname)
    colnames(betas) <- columnname
    colnames(SigEachTerm)<-columnname
    }
  }else{
    
  # Remove prefix for plotting
  columnname <- colnames(betas)
  columnname <- gsub('ple_', '', columnname)
  columnname <- gsub('_y1', '', columnname)
  colnames(betas) <- columnname
  colnames(SigEachTerm)<-columnname
  }
  
  # Melt data for plotting
  plotdf <- melt(data.frame(betas))
  # aic df
  #aicdf <- melt(data.frame(SigEachTerm))
  #colnames(aicdf)<-c('variable','Sig')
  #plotdf<-cbind(plotdf,aicdf)
  plotdf$Sig<-rep(Signifvec,each=5000)

  # and variable category labels
  categories=c('No Association','Events','Performance','Demographic','Body')
  plotdf$cat<-rep('No Association',dim(plotdf)[1])
  plotdf$cat<-factor(plotdf$cat,levels=categories)
  # Events
  plotdf$cat[plotdf$variable==c('died')]<-'Events'
  plotdf$cat[plotdf$variable==c('injured')]='Events'
  plotdf$cat[plotdf$variable==c('crime')]='Events'
  plotdf$cat[plotdf$variable==c('friend')]='Events'
  plotdf$cat[plotdf$variable==c('friend_injur')]='Events'
  plotdf$cat[plotdf$variable==c('arrest')]='Events'
  plotdf$cat[plotdf$variable==c('friend_died')]='Events'
  plotdf$cat[plotdf$variable==c('mh')]='Events'
  plotdf$cat[plotdf$variable==c('sib')]='Events'
  plotdf$cat[plotdf$variable==c('victim')]='Events'
  plotdf$cat[plotdf$variable==c('separ')]='Events'
  plotdf$cat[plotdf$variable==c('law')]='Events'
  plotdf$cat[plotdf$variable==c('school')]='Events'
  plotdf$cat[plotdf$variable==c('move')]='Events'
  plotdf$cat[plotdf$variable==c('jail')]='Events'
  plotdf$cat[plotdf$variable==c('step')]='Events'
  plotdf$cat[plotdf$variable==c('new_job')]='Events'
  plotdf$cat[plotdf$variable==c('new_sib')]='Events'

  # Performance
  plotdf$cat[plotdf$variable==c('n_g')]='Performance'
  plotdf$cat[plotdf$variable==c('n_Grades')]='Performance'
  
  # if it's a temporal precedence file, add catch 
  if (length(grep('TempPrec',file_path))>0){
  # Demographics
  plotdf$cat[plotdf$variable==c('n_interview_age')]='Demographic'
  plotdf$cat[plotdf$variable==c('income.x')]='Demographic'
  plotdf$cat[plotdf$variable==c('n_parental_education')]='Demographic'
  plotdf$cat[plotdf$variable==c('sex.xM')]='Demographic'
  plotdf$cat[plotdf$variable==c('n_race_ethnicity')]='Demographic'
  
  # swap levels of race for meaningful factor names
  levels(plotdf$variable)=c(levels(plotdf$variable),'Other','Asian','Hispanic','African American')
  plotdf$variable[plotdf$variable=='race_ethnicity.x5']='Other'
  plotdf$variable[plotdf$variable=='race_ethnicity.x4']='Asian'
  plotdf$variable[plotdf$variable=='race_ethnicity.x3']='Hispanic'
  plotdf$variable[plotdf$variable=='race_ethnicity.x2']='African American'
  }else{
  # Demographics
  plotdf$cat[plotdf$variable==c('n_interview_age')]='Demographic'
  plotdf$cat[plotdf$variable==c('n_income')]='Demographic'
  plotdf$cat[plotdf$variable==c('n_parental_education')]='Demographic'
  plotdf$cat[plotdf$variable==c('sexM')]='Demographic'
  plotdf$cat[plotdf$variable==c('n_race_ethnicity')]='Demographic'
  
  # swap levels of race for meaningful factor names
  levels(plotdf$variable)=c(levels(plotdf$variable),'Other','Asian','Hispanic','African American')
  plotdf$variable[plotdf$variable=='race_ethnicity5']='Other'
  plotdf$variable[plotdf$variable=='race_ethnicity4']='Asian'
  plotdf$variable[plotdf$variable=='race_ethnicity3']='Hispanic'
  plotdf$variable[plotdf$variable=='race_ethnicity2']='African American'
  }
  
  
  # Body
  plotdf$cat[plotdf$variabe==c('n_weight')]='Body'
  plotdf$cat[plotdf$variable==c('n_waist')]='Body'
  plotdf$cat[plotdf$variable==c('n_height')]='Body'
  plotdf$cat[plotdf$variable==c('n_BMI')]='Body'
  
  # Define custom color scheme for the categories
  colors <- c("Events" = "#FFD700", "Performance" = "#EA4335", "Demographic" = "#4285F4", "Body" = "#34A853", "No Association" = "grey")
  
  # Get levels of factor
  fct_levels <- levels(plotdf$variable)
  # Reorder levels with n_interview_age after n_Grades
  new_levels <- c(fct_levels[1:20], fct_levels[22], fct_levels[21], fct_levels[23:length(fct_levels)])
  # and n_ removal
  new_levels<- gsub("n_", "", new_levels)
  new_variable_names=gsub("n_", "", plotdf$variable)
  plotdf$variable <-factor(new_variable_names,levels=new_levels)
  # last take-outs
  plotdf=subset(plotdf, variable != "full")
  plotdf=subset(plotdf, variable != "income")
  
  # set cat insignificant to null color
  plotdf$cat[plotdf$Sig<1]='No Association'
  
 # Plot deviance explained by each term with grouping and coloration by 'cat'
  ggplot(data = plotdf, aes(x = factor(variable,levels=unique(plotdf$variable)), y = value,color=cat))  +
  geom_boxplot(width = 0.4, position = position_nudge(x = 0.5)) +
  geom_jitter(aes(color = cat), alpha = 0.01, size = 2, width = 0.2, height = 0.01) +
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  ylab("Value") +
  xlab("Covariate") +
  theme_minimal(base_size = 50) +
  ylim(c(-0.8, 0.8)) +
  labs(color = "Category")
  
  
  
  
  # ggplot(data = plotdf, aes(x = factor(variable, levels = new_levels), y = value, fill = cat))  +
  # geom_boxplot(aes(fill=cat)) +
  # scale_color_manual(values = colors) +
  # #scale_fill_manual(values = colors) +
  # ylab(label) +
  # xlab('Covariate') +
  # theme_minimal(base_size = 35) +
  # theme(axis.text.x = element_text(angle = 45)) +
  #   ylim(c(-0.005,.14))+
  #   labs(fill = "Category") 
}

```


```{r}
# load in sq
SumSq=readRDS('~/int_sumSq.rds')
# load in sq_heldout
SumSqH=readRDS('~/Desktop/g_p/int_sumSqHeldout.rds')
# load in deviance explained
DevExpl=readRDS('~/Desktop/g_p/int_DevExplained.rds')
# load in AIC
AICs=readRDS('~/Desktop/g_p/int_AIC.rds')
# get proportion of rows with lower AIC with each variable added
for (r in 1:length(colnames(AICs))){
  columnOfInt=AICs[1:35,r]
  # if removing item increases AIC (relative to full model), the item was helpful to have in the model
  propAIC=sum(columnOfInt[1:35]>AICs$AIC_full[1:35])/35
  print(colnames(AICs[r]))
  print(propAIC)
}

# visualize deviance explained by each term
# deviance explained by each term
devExplEachTerm=NULL
for (r in 1:length(colnames(DevExpl))){
  columnOfInt=DevExpl[1:35,r]
  propDev=1-(columnOfInt/DevExpl$devExplained_full[1:35])
  devExplEachTerm[r]=mean(propDev)
}

# remove prefix for plotting
columnname=colnames(DevExpl)
columnname=gsub('devExplained_','',columnname)
plotdf=data.frame(devExplEachTerm,columnname)
library(ggplot2)
# Basic barplot
ggplot(data=plotdf, aes(x=columnname, y=-devExplEachTerm)) +
  geom_bar(stat="identity")+ylab('Prop. Deviance Explained Lost Excluding Cov.')+xlab('Covariate')+theme(axis.text.x = element_text(angle=45))

```

```{r}
Int_TPDevExpl=readRDS('~/Int_TempPrec_DevExplained.rds')
# only have first 5k for now
Int_TPDevExpl=Int_TPDevExpl[1:5000,]
# visualize deviance explained by each term
# deviance explained by each term
devExplEachTerm=matrix(data=NA,nrow=5000,ncol=length(colnames(Int_TPDevExpl)))
for (r in 1:length(colnames(Int_TPDevExpl))){
  columnOfInt=Int_TPDevExpl[,r]
  propDev=1-(columnOfInt/Int_TPDevExpl$devExplained_full)
  devExplEachTerm[,r]=propDev
}

# remove prefix for plotting
columnname=colnames(Int_TPDevExpl)
columnname=gsub('devExplained_','',columnname)
colnames(devExplEachTerm)<-columnname
library(reshape2)
plotdf=melt(data.frame(devExplEachTerm))
library(ggplot2)
# Basic barplot
ggplot(data=plotdf, aes(x=variable, y=value)) +
  geom_boxplot()+ylab('Prop. Deviance Explained in TP2 Int. w/ Covariate')+xlab('Covariate')+theme(axis.text.x = element_text(angle=45))+theme_minimal()
```

```{r}
# eval AIC/BIC for temp prec
AICs=readRDS('~/p_TempPrec_AIC.rds')
AICs=AICs[1:5000,]
# get proportion of rows with lower AIC with each variable added
for (r in 1:length(colnames(AICs))){
  columnOfInt=AICs[,r]
  # if removing item increases AIC (relative to full model), the item was helpful to have in the model
  propAIC=sum(columnOfInt>AICs$AIC_full)/5000
  print(colnames(AICs[r]))
  print(propAIC)
}

```

```{r}
# sum of squares held-out temporal precedence p
int_TPDevExpl_IVE=readRDS('~/p_DevExplained.rds')
# only have first 5k for now
int_TPDevExpl_IVE=int_TPDevExpl_IVE[1:3000,]
# visualize deviance explained by each term
# deviance explained by each term
devExplEachTerm=matrix(data=NA,nrow=3000,ncol=length(colnames(int_TPDevExpl_IVE)))
for (r in 1:length(colnames(int_TPDevExpl_IVE))){
  columnOfInt=int_TPDevExpl_IVE[,r]
  propDev=1-(columnOfInt/int_TPDevExpl_IVE$devExplained_full)
  devExplEachTerm[,r]=propDev
}
# remove prefix for plotting
columnname=colnames(int_TPDevExpl_IVE)
columnname=gsub('devExplained_','',columnname)
colnames(devExplEachTerm)<-columnname
library(reshape2)
plotdf=melt(data.frame(devExplEachTerm))
library(ggplot2)
# Basic barplot
ggplot(data=plotdf, aes(x=variable, y=value)) +
  geom_boxplot()+ylab(expression('Prop. Deviance Explained: p'))+xlab('Covariate')+theme_bw(base_size=40)+theme(axis.text.x = element_text(angle=90))+ylim(c(0,1))
# comparison plot
plotdf2<-data.frame(int_TPDevExpl_IVE$devExplained_full)
colnames(plotdf2)<-'FullModel'
ggplot(data=plotdf2, aes(y=FullModel)) +
  geom_boxplot()+theme_bw(base_size=40)+ylim(c(0,1))+ylab('Deviance Explained: p')+
  scale_x_continuous(breaks = NULL)
```


```{r}
# sum of squares held-out temporal precedence p
int_TPDevExpl_IVE=readRDS('~/Int_DevExplained.rds')
# only have first 5k for now
int_TPDevExpl_IVE=int_TPDevExpl_IVE[1:3000,]
# visualize deviance explained by each term
# deviance explained by each term
devExplEachTerm=matrix(data=NA,nrow=3000,ncol=length(colnames(int_TPDevExpl_IVE)))
for (r in 1:length(colnames(int_TPDevExpl_IVE))){
  columnOfInt=int_TPDevExpl_IVE[,r]
  propDev=1-(columnOfInt/int_TPDevExpl_IVE$devExplained_full)
  devExplEachTerm[,r]=propDev
}
# remove prefix for plotting
columnname=colnames(int_TPDevExpl_IVE)
columnname=gsub('devExplained_','',columnname)
colnames(devExplEachTerm)<-columnname
library(reshape2)
plotdf=melt(data.frame(devExplEachTerm))
library(ggplot2)
# Basic barplot
ggplot(data=plotdf, aes(x=variable, y=value)) +
  geom_boxplot()+ylab(expression('Prop. Deviance Explained: Internalizing'))+xlab('Covariate')+theme_bw(base_size=40)+theme(axis.text.x = element_text(angle=90))+ylim(c(0,1))
# comparison plot
plotdf2<-data.frame(int_TPDevExpl_IVE$devExplained_full)
colnames(plotdf2)<-'FullModel'
ggplot(data=plotdf2, aes(y=FullModel)) +
  geom_boxplot()+theme_bw(base_size=40)+ylim(c(0,1))+ylab('Deviance Explained: Internalizing')+
  scale_x_continuous(breaks = NULL)
```


```{r}
# sum of squares held-out temporal precedence p
int_TPDevExpl_IVE=readRDS('~/Ext_DevExplained.rds')
# only have first 5k for now
int_TPDevExpl_IVE=int_TPDevExpl_IVE[1:3000,]
# visualize deviance explained by each term
# deviance explained by each term
devExplEachTerm=matrix(data=NA,nrow=3000,ncol=length(colnames(int_TPDevExpl_IVE)))
for (r in 1:length(colnames(int_TPDevExpl_IVE))){
  columnOfInt=int_TPDevExpl_IVE[,r]
  propDev=1-(columnOfInt/int_TPDevExpl_IVE$devExplained_full)
  devExplEachTerm[,r]=propDev
}
# remove prefix for plotting
columnname=colnames(int_TPDevExpl_IVE)
columnname=gsub('devExplained_','',columnname)
colnames(devExplEachTerm)<-columnname
library(reshape2)
plotdf=melt(data.frame(devExplEachTerm))
library(ggplot2)
# Basic barplot
ggplot(data=plotdf, aes(x=variable, y=value)) +
  geom_boxplot()+ylab(expression('Prop. Deviance Explained: Externalizing'))+xlab('Covariate')+theme_bw(base_size=40)+theme(axis.text.x = element_text(angle=90))+ylim(c(0,1))
# comparison plot
plotdf2<-data.frame(int_TPDevExpl_IVE$devExplained_full)
colnames(plotdf2)<-'FullModel'
ggplot(data=plotdf2, aes(y=FullModel)) +
  geom_boxplot()+theme_bw(base_size=40)+ylim(c(0,1))+ylab('Deviance Explained: Externalizing')+
  scale_x_continuous(breaks = NULL)
```

###### 
###### 
###### PLOT DEV EXPL BY # afflicted
###### 
###### 

```{r}
# sum of squares held-out temporal precedence p
int_TPDevExpl_IVE=readRDS('~/p_TempPrec_DevExplained.rds.rds')
NumAfflicted=readRDS('~/p_number.rds')
# correct
NumAfflicted=NumAfflicted-7686
#2-19 are relevant columns
# only have first 5k for now
int_TPDevExpl_IVE=int_TPDevExpl_IVE[1:3000,]
NumAfflicted=NumAfflicted[1:3000,]
# visualize deviance explained by each term
# deviance explained by each term
devExplEachTermPer=matrix(data=NA,nrow=3000,ncol=length(colnames(int_TPDevExpl_IVE)))
for (r in 2:19){
  columnOfInt=int_TPDevExpl_IVE[,r]
  propDev=1-(columnOfInt/int_TPDevExpl_IVE$devExplained_full)
  # r-1 because it starts on col1 where other df starts on col 2
  propDevPer=propDev/(NumAfflicted[,r-1])
  # 7,500 to guess what is would be like if everyone had it?
  devExplEachTermPer[,r]=propDevPer
}
# same rules for full-coverage terms
for (r in 20:31){
  columnOfInt=int_TPDevExpl_IVE[,r]
  propDev=1-(columnOfInt/int_TPDevExpl_IVE$devExplained_full)
  # r-1 because it starts on col1 where other df starts on col 2
  propDevPer=propDev/7686
  # 7,500 to guess what is would be like if everyone had it?
  devExplEachTermPer[,r]=propDevPer
}
# remove prefix for plotting
columnname=colnames(int_TPDevExpl_IVE)
columnname=gsub('devExplained_','',columnname)
colnames(devExplEachTermPer)<-columnname
library(reshape2)
plotdf=melt(data.frame(devExplEachTermPer))
library(ggplot2)
# Basic barplot
ggplot(data=plotdf, aes(x=variable, y=value))+ geom_boxplot()+ylab(expression('Prop. Deviance Explained Per Subj: p'))+xlab('Covariate')+theme_bw(base_size=40)+theme(axis.text.x = element_text(angle=90))+ylim(c(0,1/7686))
# comparison plot
plotdf2<-data.frame(int_TPDevExpl_IVE$devExplained_full)
colnames(plotdf2)<-'FullModel'
ggplot(data=plotdf2, aes(y=FullModel)) +
  geom_boxplot()+theme_bw(base_size=40)+ylim(c(0,1/7686))+ylab('Deviance Explained: p')+
  scale_x_continuous(breaks = NULL)
```


```{r}
# sum of squares held-out temporal precedence p
int_TPDevExpl_IVE=readRDS('~/Int_DevExplained.rds')
NumAfflicted=readRDS('~/Int_number.rds')
# correct
NumAfflicted=NumAfflicted-7686
#2-19 are relevant columns
# only have first 5k for now
int_TPDevExpl_IVE=int_TPDevExpl_IVE[1:3000,]
NumAfflicted=NumAfflicted[1:3000,]
# visualize deviance explained by each term
# deviance explained by each term
devExplEachTermPer=matrix(data=NA,nrow=3000,ncol=length(colnames(int_TPDevExpl_IVE)))
for (r in 2:19){
  columnOfInt=int_TPDevExpl_IVE[,r]
  propDev=1-(columnOfInt/int_TPDevExpl_IVE$devExplained_full)
  # r-1 because it starts on col1 where other df starts on col 2
  propDevPer=propDev/(NumAfflicted[,r-1])
  # 7,500 to guess what is would be like if everyone had it?
  devExplEachTermPer[,r]=propDevPer
}
# same rules for full-coverage terms
for (r in 20:31){
  columnOfInt=int_TPDevExpl_IVE[,r]
  propDev=1-(columnOfInt/int_TPDevExpl_IVE$devExplained_full)
  # r-1 because it starts on col1 where other df starts on col 2
  propDevPer=propDev/7686
  # 7,500 to guess what is would be like if everyone had it?
  devExplEachTermPer[,r]=propDevPer
}
# remove prefix for plotting
columnname=colnames(int_TPDevExpl_IVE)
columnname=gsub('devExplained_','',columnname)
colnames(devExplEachTermPer)<-columnname
library(reshape2)
plotdf=melt(data.frame(devExplEachTermPer))
library(ggplot2)
# Basic barplot
ggplot(data=plotdf, aes(x=variable, y=value))+ geom_boxplot()+ylab(expression('Prop. Deviance Explained Per Subj: Internalizing'))+xlab('Covariate')+theme_bw(base_size=40)+theme(axis.text.x = element_text(angle=90))+ylim(c(0,1/7686))
# comparison plot
plotdf2<-data.frame(int_TPDevExpl_IVE$devExplained_full)
colnames(plotdf2)<-'FullModel'
ggplot(data=plotdf2, aes(y=FullModel)) +
  geom_boxplot()+theme_bw(base_size=40)+ylim(c(0,1/7686))+ylab('Deviance Explained: Internalizing')+
  scale_x_continuous(breaks = NULL)
```

```{r}
# sum of squares held-out temporal precedence p
int_TPDevExpl_IVE=readRDS('~/Int_DevExplained.rds')
NumAfflicted=readRDS('~/Ext_number.rds')
# correct
NumAfflicted=NumAfflicted-7686
#2-19 are relevant columns
# only have first 5k for now
int_TPDevExpl_IVE=int_TPDevExpl_IVE[1:3000,]
NumAfflicted=NumAfflicted[1:3000,]
# visualize deviance explained by each term
# deviance explained by each term
devExplEachTermPer=matrix(data=NA,nrow=3000,ncol=length(colnames(int_TPDevExpl_IVE)))
for (r in 2:19){
  columnOfInt=int_TPDevExpl_IVE[,r]
  propDev=1-(columnOfInt/int_TPDevExpl_IVE$devExplained_full)
  # r-1 because it starts on col1 where other df starts on col 2
  propDevPer=propDev/(NumAfflicted[,r-1])
  # 7,500 to guess what is would be like if everyone had it?
  devExplEachTermPer[,r]=propDevPer
}
# same rules for full-coverage terms
for (r in 20:31){
  columnOfInt=int_TPDevExpl_IVE[,r]
  propDev=1-(columnOfInt/int_TPDevExpl_IVE$devExplained_full)
  # r-1 because it starts on col1 where other df starts on col 2
  propDevPer=propDev/7686
  # 7,500 to guess what is would be like if everyone had it?
  devExplEachTermPer[,r]=propDevPer
}
# remove prefix for plotting
columnname=colnames(int_TPDevExpl_IVE)
columnname=gsub('devExplained_','',columnname)
colnames(devExplEachTermPer)<-columnname
library(reshape2)
plotdf=melt(data.frame(devExplEachTermPer))
library(ggplot2)
# Basic barplot
ggplot(data=plotdf, aes(x=variable, y=value))+ geom_boxplot()+ylab(expression('Prop. Deviance Explained Per Subj: Externalizing'))+xlab('Covariate')+theme_bw(base_size=40)+theme(axis.text.x = element_text(angle=90))+ylim(c(0,1/7686))
# comparison plot
plotdf2<-data.frame(int_TPDevExpl_IVE$devExplained_full)
colnames(plotdf2)<-'FullModel'
ggplot(data=plotdf2, aes(y=FullModel)) +
  geom_boxplot()+theme_bw(base_size=40)+ylab('Deviance Explained: Internalizing')+
  scale_x_continuous(breaks = NULL)+ylim(c(0,.5))
```

```{r}
# sum of squares held-out temporal precedence p
int_TPDevExpl_IVE=readRDS('~/p_DevExplained.rds')
# only have first 5k for now
int_TPDevExpl_IVE=int_TPDevExpl_IVE[1:3000,]
# visualize deviance explained by each term
# deviance explained by each term
devExplEachTerm=matrix(data=NA,nrow=3000,ncol=length(colnames(int_TPDevExpl_IVE)))
for (r in 1:length(colnames(int_TPDevExpl_IVE))){
  columnOfInt=int_TPDevExpl_IVE[,r]
  propDev=1-(columnOfInt/int_TPDevExpl_IVE$devExplained_full)
  devExplEachTerm[,r]=propDev
}
# remove prefix for plotting
columnname=colnames(int_TPDevExpl_IVE)
columnname=gsub('devExplained_','',columnname)
colnames(devExplEachTerm)<-columnname
library(reshape2)
plotdf=melt(data.frame(devExplEachTerm))
library(ggplot2)
# Basic barplot
ggplot(data=plotdf, aes(x=variable, y=value)) +
  geom_boxplot()+ylab(expression('Prop. Deviance Explained: p'))+xlab('Covariate')+theme_bw(base_size=40)+theme(axis.text.x = element_text(angle=90))+ylim(c(0,1))
# comparison plot
plotdf2<-data.frame(int_TPDevExpl_IVE$devExplained_full)
colnames(plotdf2)<-'FullModel'
ggplot(data=plotdf2, aes(y=FullModel)) +
  geom_boxplot()+theme_bw(base_size=40)+ylim(c(0,1))+ylab('Deviance Explained: p')+
  scale_x_continuous(breaks = NULL)
```


```{r}
# sum of squares held-out temporal precedence p
int_TPDevExpl_IVE=readRDS('~/DevExpl/Ext_TempPrec_IVE_DevExplained.rds')
# only have first 5k for now
int_TPDevExpl_IVE=int_TPDevExpl_IVE[1:5000,]
# visualize deviance explained by each term
# deviance explained by each term
devExplEachTerm=matrix(data=NA,nrow=5000,ncol=length(colnames(int_TPDevExpl_IVE)))
for (r in 1:length(colnames(int_TPDevExpl_IVE))){
  columnOfInt=int_TPDevExpl_IVE[,r]
  propDev=1-(columnOfInt/int_TPDevExpl_IVE$devExplained_full)
  devExplEachTerm[,r]=propDev
}
# remove prefix for plotting
columnname=colnames(int_TPDevExpl_IVE)
columnname=gsub('devExplained_','',columnname)
colnames(devExplEachTerm)<-columnname
library(reshape2)
plotdf=melt(data.frame(devExplEachTerm))
library(ggplot2)
# Basic barplot
ggplot(data=plotdf, aes(x=variable, y=value)) +
  geom_boxplot()+ylab(expression('Prop. Deviance Explained TP2: p'))+xlab('Covariate')+theme_bw(base_size=40)+theme(axis.text.x = element_text(angle=90))+ylim(c(0,1))
# comparison plot
plotdf2<-data.frame(int_TPDevExpl_IVE$devExplained_full)
colnames(plotdf2)<-'FullModel'

ggplot(aes(y=FullModel),data=plotdf2)+geom_boxplot()+theme_minimal(base_size=40)+ylab('Deviance Explained')+ylim(c(.23,.65))


```

```{r}
# intervisit events versus happened before baseline
yleCols=colnames(int_TPDevExpl_IVE)[2:19]
b4=Int_TPDevExpl[,c(yleCols)]
# normalize to variance explained in that boot
for (r in 1:5000){
  b4[r,]=b4[r,]/ Int_TPDevExpl$devExplained_full[r]
}
# now for ive
IVE=int_TPDevExpl_IVE[,c(yleCols)]
for (r in 1:5000){
  IVE[r,]=IVE[r,]/ int_TPDevExpl_IVE$devExplained_full[r]
}
mb4=melt(b4)
mb4$when='Before'
mIve=melt(IVE)
mIve$when='After'
merged=rbind(mb4,mIve)

# plot difference in deviance explained
ggplot(data=merged, aes(x=variable, y=value, color=when)) +
  geom_boxplot()+ylab('Deviance Explained in TP2 Int. w/ event occuring')+xlab('Covariate')+theme(axis.text.x = element_text(angle=45))+theme_minimal()
```

```{r}
min_one_exp <- function(x) {
        exp(x - 1)
    }
```

```{r}
masterdf$Grades<-as.ordered(masterdf$Grades)

# plot g with grades incldued

fullModel<-bam(cbcl_scr_syn_totprob_r~ple_died_y+ple_injured_y+ple_crime_y+ple_friend_y+ple_friend_injur_y+ple_arrest_y+ple_friend_died_y+ple_mh_y+ple_sib_y+ple_victim_y+ple_separ_y+ple_law_y+ple_school_y+ple_move_y+ple_jail_y+ple_step_y+ple_new_job_y+ple_new_sib_y+s(g,k=4)+s(interview_age,k=4)+s(parentPcount,k=4)+income+s(parental_education,k=4)+sex+race_ethnicity+s(weight,k=4)+s(waist,k=4)+s(height,k=4)+s(BMI,k=4)+GradesO,data=masterdf)

visreg(fullModel,"g",gg=TRUE,ylab="Total Symptoms (Residualized)",xlab="g (Grades accounted for)",line.par=list(size=2,col='black'),alpha=0.05)+theme_minimal(base_size=25)+ylim(c(-.2,6.2))

visreg(fullModel,"Grades",gg=TRUE,ylab="Total Symptoms (Residualized)",line.par=list(size=2,col='black'),alpha=0.05)+theme_minimal(base_size=25)+ylim(c(-.2,6.2))


# plot g without grades

fullModel<-bam(cbcl_scr_syn_totprob_r~ple_died_y+ple_injured_y+ple_crime_y+ple_friend_y+ple_friend_injur_y+ple_arrest_y+ple_friend_died_y+ple_mh_y+ple_sib_y+ple_victim_y+ple_separ_y+ple_law_y+ple_school_y+ple_move_y+ple_jail_y+ple_step_y+ple_new_job_y+ple_new_sib_y+s(g,k=4)+s(interview_age,k=4)+s(parentPcount,k=4)+income+s(parental_education,k=4)+sex+race_ethnicity+s(weight,k=4)+s(waist,k=4)+s(height,k=4)+s(BMI,k=4),data=masterdf)

visreg(fullModel,"g",gg=TRUE,ylab="Total Symptoms (Residualized)",line.par=list(size=2,col='black'),alpha=0.05)+theme_minimal(base_size=25)+ylim(c(-.2,6.2))
```


```{r}
# p
fullModel<-bam(cbcl_scr_syn_totprob_r~ple_died_y+ple_injured_y+ple_crime_y+ple_friend_y+ple_friend_injur_y+ple_arrest_y+ple_friend_died_y+ple_mh_y+ple_sib_y+ple_victim_y+ple_separ_y+ple_law_y+ple_school_y+ple_move_y+ple_jail_y+ple_step_y+ple_new_job_y+ple_new_sib_y+s(g,k=4)+s(interview_age,k=4)+s(parentPcount,k=4)+income+s(parental_education,k=4)+sex+race_ethnicity+s(weight,k=4)+s(waist,k=4)+s(height,k=4)+s(BMI,k=4)+GradesO,data=masterdf)
a<-summary(fullModel)
SignifvecP=rep(0,24)
SignifvecP[a$p.pv[2:25]<0.01]=1
plot_betas('~/Betas/p_betas.rds',SignifvecP)
# internalizing
fullModel<-bam(cbcl_scr_syn_internal_r~ple_died_y+ple_injured_y+ple_crime_y+ple_friend_y+ple_friend_injur_y+ple_arrest_y+ple_friend_died_y+ple_mh_y+ple_sib_y+ple_victim_y+ple_separ_y+ple_law_y+ple_school_y+ple_move_y+ple_jail_y+ple_step_y+ple_new_job_y+ple_new_sib_y+s(g,k=4)+s(interview_age,k=4)+s(parentPcount,k=4)+income+s(parental_education,k=4)+sex+race_ethnicity+s(weight,k=4)+s(waist,k=4)+s(height,k=4)+s(BMI,k=4)+GradesO,data=masterdf)
a<-summary(fullModel)
SignifvecI=rep(0,24)
SignifvecI[a$p.pv[2:25]<0.01]=1
plot_betas('~/Betas/Int_betas.rds',SignifvecI)
# externalizing
fullModel<-bam(cbcl_scr_syn_external_r~ple_died_y+ple_injured_y+ple_crime_y+ple_friend_y+ple_friend_injur_y+ple_arrest_y+ple_friend_died_y+ple_mh_y+ple_sib_y+ple_victim_y+ple_separ_y+ple_law_y+ple_school_y+ple_move_y+ple_jail_y+ple_step_y+ple_new_job_y+ple_new_sib_y+s(g,k=4)+s(interview_age,k=4)+s(parentPcount,k=4)+income+s(parental_education,k=4)+sex+race_ethnicity+s(weight,k=4)+s(waist,k=4)+s(height,k=4)+s(BMI,k=4)+GradesO,data=masterdf)
a<-summary(fullModel)
SignifvecE=rep(0,24)
SignifvecE[a$p.pv[2:25]<0.01]=1
plot_betas('~/Betas/Ext_betas.rds',Signifvec)
```

```{r}
# plot deviance explained with extra terms
Pmodel<-bam(cbcl_scr_syn_totprob_r.y~cbcl_scr_syn_totprob_r.x,data=masterdf)
nullDev=summary(Pmodel)$dev.expl
Pdevs=readRDS('~/DevExpl/p_TempPrec_IVE_DevExplained.rds')[1:5000,]
# Create histogram
p <- ggplot(data = data.frame(x = Pdevs[,1]), aes(x = x)) + 
    geom_histogram(aes(y = ..density..), bins = 30, fill = "white", color = "black")

# Add vertical line at nullDev
p + geom_vline(xintercept = nullDev, color = "#4285F4", size = 1)+theme_minimal(base_size=30)+xlab('Deviance explained with all other measures: p')
```


# plot grades

# Predict values and standard errors
df <- data.frame(Grades = factor(1:5), 
                 preds = predict(fullModel, newdata = data.frame(Grades = factor(1:5)), se.fit = TRUE)$fit, 
                 se = predict(fullModel, newdata = data.frame(Grades = factor(1:5)), se.fit = TRUE)$se.fit)





masterdf$preds <- predict(fullModel, masterdf)
masterdf$Grades <- factor(masterdf$Grades, levels = unique(masterdf$Grades))

# Calculate standard errors for predicted values
se <- sqrt(diag(vcov(fullModel)))

# Plot predicted values as horizontal lines
ggplot(masterdf, aes(x = as.numeric(Grades), y = min_one_exp(preds))) +
  geom_pointrange(aes(ymin = lower, ymax = upper))
 #geom_errorbar(aes(ymin = preds - se, ymax = preds + se), width = 0.2, color = "#2C3E50", size = 1) +
  geom_point(aes(y = 'cbcl_scr_syn_totprob_r'), alpha = 0.5, color = "#A3B56B", size = 3) +
  geom_line(color = "#2C3E50", size = 1) +
  ylim(c(0, 40)) +
  ylab("Total Symptoms") +
  theme_minimal(base_size = 25)






# Get predicted values from model
preds <- predict(fullModel, type = "response")

# Create a data frame with residuals and predicted values
df <- data.frame(Grades = factor(masterdf$Grades, levels = 1:5), 
                 preds = preds, 
                 reside = residuals(fullModel, type = "response"))

# Order the data by Grades
df <- df[order(df$Grades), ]

f# Create the plot
ggplot(df, aes(x = Grades, y = preds)) +
  geom_hline(yintercept = (df$preds + df$reside)) +
  scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
  ylim(c(0, 40)) +
  ylab("Total Symptoms") +
  theme_minimal(base_size = 25)

```

```{r}
library(ggplot2)

# Create a data frame with predicted values and residuals
masterdf$predicted=predict(fullModel, type = "response")
masterdf$resids=residuals(fullModel)

# Extract the coefficients from the model
coefs <- coef(fullModel)
intcpt <- coefs[1]
Grades_coefs <- coefs[2:length(coefs)]

# Create a sequence of values for x-axis
Grades_seq<-c(1,2,3,4,5)

# Create a data frame with predicted values for each factor level
pred_df <- data.frame(Grades = Grades_seq)
for (i in 1:length(Grades_coefs)) {
  pred_df[[paste0("pred_", i)]] <- predict(fullModel, newdata = data.frame(Grades = factor(Grades_seq, levels = levels(masterdf$Grades))), type = "response") - intcpt + Grades_coefs[i]
}

# Convert data to long format for plotting
pred_df_long <- reshape2::melt(pred_df, id.vars = "Grades", variable.name = "group", value.name = "preds")

# Add standard errors to predicted values
pred_df_long$se <- summary(fullModel)$sigma

# Plot predicted values with standard errors and residuals
ggplot(masterdf, aes(x = Grades, y = preds)) +
  geom_point(alpha = 0.5, size = 3, aes(color = resids)) +
  geom_errorbar(data = pred_df_long, aes(x = Grades, y = preds, ymin = preds - se, ymax = preds + se, group = group), width = 0.1, color = "#2C3E50", size = 2) +
  scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
  ylim(c(0, 40)) +
  ylab("Total Symptoms") +
  theme_minimal(base_size = 25)

```

masterdf
fullModel<-bam(cbcl_scr_syn_totprob_r~ple_died_y+ple_injured_y+ple_crime_y+ple_friend_y+ple_friend_injur_y+ple_arrest_y+ple_friend_died_y+ple_mh_y+ple_sib_y+ple_victim_y+ple_separ_y+ple_law_y+ple_school_y+ple_move_y+ple_jail_y+ple_step_y+ple_new_job_y+ple_new_sib_y+s(g,k=4)+s(interview_age,k=4)+s(parentPcount,k=4)+income+s(parental_education,k=4)+sex+race_ethnicity+s(weight,k=4)+s(waist,k=4)+s(height,k=4)+s(BMI,k=4)+Grades,data=masterdf)

visreg(fullModel,"Grades",gg=TRUE,ylab="Total Symptoms",trans=min_one_exp,rug=FALSE,line.par=list(size=2,col='black'))+geom_point(alpha=0.05,size=3,color='#1d6628')+theme_minimal(base_size=25)+ylim(c(0,40))

fullModel<-bam(cbcl_scr_syn_totprob_r~ple_died_y+ple_injured_y+ple_crime_y+ple_friend_y+ple_friend_injur_y+ple_arrest_y+ple_friend_died_y+ple_mh_y+ple_sib_y+ple_victim_y+ple_separ_y+ple_law_y+ple_school_y+ple_move_y+ple_jail_y+ple_step_y+ple_new_job_y+ple_new_sib_y+s(g,k=4)+s(interview_age,k=4)+s(parentPcount,k=4)+income+s(parental_education,k=4)+sex+race_ethnicity+s(weight,k=4)+s(waist,k=4)+s(height,k=4)+s(BMI,k=4),data=masterdf)

visreg(fullModel,"g",gg=TRUE,ylab="Total Symptoms",trans=min_one_exp,rug=FALSE,line.par=list(size=2,col='black'))+geom_point(alpha=0.05,size=3,color='#1d6628')+theme_minimal(base_size=25)+ylim(c(0,40))


```